---
title: "PUBHLTH 460: Final project individual report"
author: "Elaine Xu"
fontsize: 11pt
date: "11/19/2020"
output: github_document
---

```{r, include=FALSE}
library(stringr) #assist with text manipulation
library(dplyr) # data manipulation
library(readr) # data input
library(caret) #select tuning parameters
library(MASS) # contains the data
library(nnet) # used for Multinomial Classification 
library(readr) #assist with text manipulation
library(kernlab) #assist with SVM feature selection
library(class) # used for an object-oriented style of programmin
library(KernelKnn) # used for K- Nearest-Neighbors method
library(nnet) # Used for Neural Net
library(e1071) 
library(gbm)
library(xgboost) # Used for xgbTree
library(ggcorrplot)
```

For this project, my team mates and I were trying to find out if there exists a relationship between protein somatic mutation and cancer type, and the proteins that are relatively significant. We were using linear regression, heat map and Bonferroni correction to approach our goal. We used two datasets, which are 77_cancer_proteomes_CPTAC_itraq and clinical_data_breast_cancer. 

```{r,eval=FALSE}
protein <-read.csv("77_cancer_proteomes_CPTAC_itraq.csv")
dim(protein) #86 records, 12553 predictors
clin <-read.csv("clinical_data_breast_cancer.csv")
dim(clin) #105 Records, 30 clinical predictors/measurements
```

By observing the two datasets, we found out that the first column of the clinical dataset and the first row of the proteomics dataset are both represent the patient ID. Since the protein and patient names are not formatted exactly the same as needed for downstream analysis in two datasets. 

We first rewrite the patient ID in 77_cancer dataset into the same form as the patient ID in clinical dataset. To combine the two datasets, first we rotate the 77_cancer dataset table. We want the patient ID to be the first column, because it will make it easier to combine the two datasets. However, we didn't find a way to do this in R, so we just rotate the table by copy the original dataset and paste it by using the "transform" into a new excel form. By changing the patient ID column in both datasets from factor to character, we joined the two table.

```{r,eval=FALSE}
colnames(protein[0,4:ncol(protein)])
#manipulate dataframe
p1<-str_replace(colnames(protein[4:ncol(protein)]),"TCGA","")
p2<-str_replace(p1,".\\d+$","")
p3<-str_replace(p2,"^","TCGA-")
p4 = str_replace(p3,"\\.","-")
names(protein)[4:ncol(protein)] = p4
protein = protein[,!duplicated(colnames(protein))]

#edit column name 
names(protein)[1] = "Complete.TCGA.ID"

#save all 
np<-as.character(protein[,1])

#trim gen column
protein<-protein[,c(1,4:83)]
```

```{r,eval=FALSE}

protein.t.test <-read.csv("protein.t_test.csv")

#change the primary key from factor to character
clin$Complete.TCGA.ID<-as.character(clin$Complete.TCGA.ID)
protein.t.test$Complete.TCGA.ID<-as.character(protein.t.test$Complete.TCGA.ID)
#join two table
MASTER.test<- inner_join(protein.t.test,clin, by="Complete.TCGA.ID")
#export data
write.csv(MASTER.test,"MASTER.test.csv", row.names = TRUE)
```


```{r,eval=FALSE}
MASTER <-read.csv("MASTER.test.csv")
#y is MASTER$PAM50.mRNA
#x is all np variables

#factor y 
MASTER$PAM50.mRNA<-as.factor(MASTER$PAM50.mRNA)
levels(MASTER$PAM50.mRNA)

MASTER.try<-MASTER
#delete na columns
MASTER.try.omitna.df<-MASTER.try[ , colSums(is.na(MASTER.try)) == 0]
```

In the dataset, there has several missing data and some text data, such as patient's gender. This makes the data hard to be analyzed, so we decide only use the data of level of protein which are all numbers and omit the proteins that have missing data. After cleaning the data, since we only want to analysis luminal A and luminal B, so we only leave rows that in the PAM50 column are luminal A or B.

```{r,eval=FALSE}
#leave only rows where PAM50.mRNA==lumin A B
MASTER.try.omitna.A.df<-MASTER.try.omitna.df[MASTER.try.omitna.df$PAM50.mRNA=="Luminal A",]
MASTER.try.omitna.B.df<-MASTER.try.omitna.df[MASTER.try.omitna.df$PAM50.mRNA=="Luminal B",]
MASTER.try.omitna.AB.df<-rbind(MASTER.try.omitna.A.df,MASTER.try.omitna.B.df)
MASTER.try.omitna.AB.df
```

